Fusion Moves for Correlation Clustering

### 2014-11-12 paper/egpaper_for_review.tex
update sec 1-4

% Include other packages here, before hyperref.
\usepackage{my_macros}
\usepackage{paralist}
\usepackage{amsthm}
% \usepackage{dirtytalk} fah: my miktex didn't like this; and we don't need it here. 
%\usepackage{framed}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{setspace}
%for tikz
\usepackage{tikz,pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
%\usepackage{scrextend}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\pdfpageattr{/Group <</S /Transparency /I true /CS /DeviceRGB>>}

@@

\input{inputs/fig_hc_alg.tex}
\section{Energy Based Hierarchical Clustering}\label{sec:ehc}
Agglomerative hierarchical clustering (HC) is widely used
in graph / image segmentation~\cite{arbelaez_2006}.
In each step, the edge with the highest weight $w$ is
contracted (red edges in Fig.~\ref{fig:hc_alg}).
Doing so, parallel edges edges can occur.
In agglomerative clustering, weights of
parallel edges are merged into single edges.
For image segmentation, the length
weighted mean is used to do this update~\cite{arbelaez_2006}.

While we, contrary to~\cite{arbelaez_2006}, directly work on energies, we are interesting in an 
agglomeration with respect to our energy function.
For this we propose to use energy based agglomeration
withe the following update rule:
Whenever there are multiple edges between 
a pair of nodes, these edges are merged into a single edge
by summing up their weights, since we 
minimize the sum of the cut edges.
We call HC with this update method Energy base Hierarchical Clustering (EHC).

We stop EHC if the highest edge weight is
smaller or equal to zero  (green edge in Fig.~\ref{fig:hc_alg}). Any further edge
contraction leads to worse energies.

@@

Given the intrinsic greediness of hierarchical clustering, 
and the fact that EHC does only local decisions, it is 
sensitive against salt and paper noise 
and we do not expect very good solutions.

@@

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning (see Fig.~\ref{fig:fusion_abiguty}).
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}    & =  y' \cup y'' \\
\mathcal{C}  & =  \{ (i, j) \quad | \quad \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{C}$.
%
\begin{align}
  y^* &=& \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
  s.t & &  y_{ij} = 0 \quad \forall (i, j) \in \mathcal{C} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy:
\begin{align}
  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solutions.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.
%
How to optimize  Eq.~\ref{eq:fusion_move_a} in an more efficient way is
described in the following section.

@@

\begin{table}[H]
   \tiny
   \centering
   \caption{Evaluation of Variation of Information (VI) for datasets with available ground-truth.}
   \label{tab:eval}
   \begin{tabular}{lllllll}
      \toprule
         Dataset          &  image-seg   & knott-3d-150&knott-3d-300&knott-3d-450 &knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &   $ 4.9633$  &    $ 2.9936$ &       --       &\\ 
         HC               &   $ 2.5967$  &    $ 1.5477$ &    $ 2.3513$   &    $ 2.9155$ \\
         HC-CGC           &   $ 2.5164$  &    $ 0.9052$ &    $ 1.7636$   &    $ 2.2256$ \\
         CGC              &   $ 2.5247$  &    $ 0.9267$ &    $ 1.8822$   &    $ 2.3104$\\
         KL               &   $ 2.6432$  &    $ 2.0648$ &    $ 4.1318$   &    $ 4.9270$\\
         FUSION           &      --      &    $ 2.8787$ &    $ 4.0744$   &    $ 4.6616$\\
         MC-R             &   $ 2.5471$  &    $ 0.9178$ &    $ 1.6369$   &    $ 2.8710$\\   
         MC-I             &   $ 2.5367$  & $\bf 0.9063$ & $\bf 1.6352$   & $\bf 2.0037$ \\  
         FusionCC-HC-MC   &   $ 2.5319$  &    $ 0.9629$ &    $ 1.6516$   &    $ 2.0801$\\  
         FusionCC-HC-CGC  &$\bf 2.4961$  &    $ 0.9679$ &    $ 1.7673$   &    $ 2.3809$ \\  
         FusionCC-WS-MC   &   $ 2.5340$  &    $ 0.9629$ &    $ 1.6742$   &    $ 2.0739$\\  
         FusionCC-WS-CGC  &   $ 2.5192$  &    $ 1.0585$ &    $ 2.1344$   &    $ 2.7487$ \\      
      \bottomrule
   \end{tabular}
\end{table}

@@

We have presented a fast and scalable 
approximate solver for correlation 
clustering, named Fusion Moves for correlation clustering (FM-CC).
The best solution is iteratively improved 
by a fusion with proposal solutions.
The itself is formulated as correlation
clustering on a smaller graph fewer edges and nodes
and can therefore be solved very fast.
Our evaluation shows that CC-FF
significantly outperforms 
other solvers w.r.t. any time performance 
with increasing problem size.

>>>>

% Include other packages here, before hyperref.
\usepackage{my_macros}
\usepackage{paralist}
\usepackage{amsthm}
% \usepackage{dirtytalk} fah: my miktex didn't like this; and we don't need it here. 
%\usepackage{framed}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{setspace}
%for tikz
\usepackage{tikz,pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
%\usepackage{scrextend}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\pdfpageattr{/Group <</S /Transparency /I true /CS /DeviceRGB>>}
\usepackage{everypage}
\AddEverypageHook{%
  \makeatletter%
  \special{pdf: put @thispage <</Group << /S /Transparency /I true /CS /DeviceRGB>> >>}%
  \makeatother%
}%

@@

\input{inputs/fig_hc_alg.tex}
\section{Energy Based Hierarchical Clustering}\label{sec:ehc}
Agglomerative hierarchical clustering (HC) is widely used
in graph / image segmentation~\cite{arbelaez_2006}.
In each step, the edge with the highest weight $w$ is
contracted (green edges in Fig.~\ref{fig:hc_alg}).
Doing so, parallel edges edges can occur.
In agglomerative clustering, weights of
parallel edges are merged into single edges.
For image segmentation, the length
weighted mean is used to do this update~\cite{arbelaez_2006}.

While we, contrary to~\cite{arbelaez_2006}, directly work on energies, we are interesting in an 
agglomeration with respect to our energy function~\ref{eq:edgeproblem}.
For this we propose to use energy based agglomeration
with the following update rule:
Whenever there are multiple edges between 
a pair of nodes, these edges are merged into a single edge
by summing up their weights, since we 
minimize the sum of the cut edges.
We call HC with this update method Energy base Hierarchical Clustering (EHC).

We stop EHC if the highest edge weight is
smaller or equal to zero  (blue edge in Fig.~\ref{fig:hc_alg}). Any further edge
contraction leads to worse energies.

@@

Given the intrinsic greediness of hierarchical clustering, 
and the fact that EHC does only local decisions, it is 
sensitive against noise and we cannot expect optimal solutions.

@@

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning, \cf Fig.~\ref{fig:fusion_abiguty}.
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}_{ij}    & = \max\{ y_{ij}', y_{ij}''\}  & \forall {ij}\in E\\  % y' \cup y'' \\
\mathcal{E}_0  & =  \{ (i, j) \; | \; \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{E}_0$.
%
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
      &\quad \textrm{s.t.} \quad y_{ij} = 0 & \forall (i, j) \in \mathcal{E}_0 \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy,
because $y'$ and $y''$ are feasible solutions for problem~\ref{eq:fusion_move_a}.
%\begin{align}
%  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
%\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solution.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.
%
How to optimize  Eq.~\ref{eq:fusion_move_a} in an more efficient way is
described in the following section.

@@

\begin{table}[H]
   \tiny
   \centering
   \caption{Evaluation of Variation of Information (VI) for datasets with available ground-truth.}
   \label{tab:eval}
   \begin{tabular}{lllllll}
      \toprule
         Dataset          &  image-seg   & knott-3d-150 & knott-3d-300   & knott-3d-450 & knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &   $ 4.9633$  &    $ 2.9936$ &       --       &              &              &  --\\ 
         HC               &   $ 2.5967$  &    $ 1.5477$ &    $ 2.3513$   &    $ 2.9155$ &              & $       2.8395$\\
         HC-CGC           &   $ 2.5164$  &    $ 0.9052$ &    $ 1.7636$   &    $ 2.2256$ &              & $       1.7603$ \\
         CGC              &   $ 2.5247$  &    $ 0.9267$ &    $ 1.8822$   &    $ 2.3104$ &              & $       6.8908$ \\
         KL               &   $ 2.6432$  &    $ 2.0648$ &    $ 4.1318$   &    $ 4.9270$ &              & $       7.1057$\\
         FUSION           &      --      &    $ 2.8787$ &    $ 4.0744$   &    $ 4.6616$ &              & $       6.5366$ \\
         MC-R             &   $ 2.5471$  &    $ 0.9178$ &    $ 1.6369$   &    $ 2.8710$ &              & $       6.5058$\\   
         MC-I             &   $ 2.5367$  & $\bf 0.9063$ & $\bf 1.6352$   & $\bf 2.0037$ &              & $       4.3319$\\  
         FusionCC-HC-MC   &   $ 2.5319$  &    $ 0.9629$ &    $ 1.6516$   &    $ 2.0801$ &              & $\bf    1.3347$ \\  
         FusionCC-HC-CGC  &$\bf 2.4961$  &    $ 0.9679$ &    $ 1.7673$   &    $ 2.3809$ &              & $       2.1347$ \\  
         FusionCC-WS-MC   &   $ 2.5340$  &    $ 0.9629$ &    $ 1.6742$   &    $ 2.0739$ &              & $\bf    1.3334$ \\  
         FusionCC-WS-CGC  &   $ 2.5192$  &    $ 1.0585$ &    $ 2.1344$   &    $ 2.7487$ &              & $       3.3514$ \\      
      \bottomrule
   \end{tabular}
\end{table}

@@

We have presented a fast and scalable 
approximate solver for correlation 
clustering, named Fusion Moves for correlation clustering (FM-CC).
The best solution is iteratively improved 
by a fusion with proposal solutions.
The fusion move itself is formulated as correlation
clustering on a smaller graph fewer edges and nodes
and can therefore be solved very fast.
Our evaluation shows that CC-FF
significantly outperforms 
other solvers w.r.t. any time performance 
with increasing problem size.

### 2014-11-12 paper/introduction-new.tex
update sec 1-4

\vspace{0.1cm}
\noindent \textbf{Outline:} 
In sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
In sec.~\ref{sec:related_work} we will 
discuss existing methods for correlation 
clustering and briefly explain the concept of fusion moves.
In sec: ~\ref{sec:cc_fm} we describe our proposed
method and show the properties of the algorithm.
In sec.~\ref{sec:exp} we show an evaluation
of the method on existing parameter  and discuss the effects of parameters.
Future work will be discussed in sec. \ref{sec:future} and
we will conclude in ~\ref{sec:conclusion}.
%-------------------------------------------------------------------------

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
%A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
%of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{kappes_2013_arxiv}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $v$ a label $l_v \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align} 
or in the edge domain, by label each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{kappes_2013_arxiv} both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the formulation~\cite{kappes_2011_emmcvpr}.

>>>>

\vspace{0.1cm}
\noindent \textbf{Outline:} 
In Sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
In Sec.~\ref{sec:cc_fm} we describe our proposed
method and show the properties of the algorithm.
In Sec.~\ref{sec:exp} we show an evaluation
of the proposed methods methods and state of the art
and conclude in Sec.~\ref{sec:future} and\ref{sec:conclusion}.
%-------------------------------------------------------------------------

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
We will use $w_e$ as a shorthand for $w(e)$.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
%A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
%of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$, \cf Fig.~\ref{fig:notation} .
An edge labeling is only consistent if it contains no dangling edges~\cite{kappes_2013_arxiv}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $v$ a label $l_v \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align} 
or in the edge domain, by label each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{kappes_2013_arxiv}  both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the formulation, \cf Fig.~\ref{fig:notation}.
